### 2023-07-02 Article.tex
First upload
This is the first upload

%% Author_tex.tex
%% V1.0
%% 2016/09/02
%% developed by Techset
%%
%% This file describes the coding for cta-author.cls

\documentclass{cta-author}%%%%where cta-author is the template name

\newtheorem{theorem}{Theorem}{}
\newtheorem{corollary}{Corollary}{}
\newtheorem{remark}{Remark}{}

%%%%%%Begin MyPackages%%
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{comment}
\usepackage{framed}
\usepackage{multirow}
\usepackage{color}
\usepackage{tabulary}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{url}
\usepackage{soul}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{caption}
\usepackage[title]{appendix}
\usepackage{ulem}
\usepackage{placeins}
\usepackage{balance}
\usepackage[autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
%%%%%%end MyPackages%%%%

\begin{document}

\title{How does an Experimental Software Engineering Lab Work? An Ethnographic Study}

\author{\au{Efraín R. Fonseca C.$^{1\corr}$}, \au{Oscar Dieste$^{2}$}, \au{Natalia Juristo$^{2}$}}

\address{\add{1}{Departamento de Ciencias de la Computación, Universidad de las Fuerzas Armadas - ESPE, Av. General Rumiñahui s/n y Ambato, Quito, Ecuador}
\add{2}{Departamento de Lenguajes y Sistemas Informáticos e Ingeniería de Software, Universidad Politécnica de Madrid, Calle de los Ciruelos, 28660, Madrid, España}
%\add{3}{Third Department, Third University, Address, Country Name}
%\add{4}{Current affiliation: Fourth Department, Fourth University, Address, Country Name}
\email{erfonseca@espe.edu.ec}}

\input{tex/abstract}

\maketitle

%%%%%%% Begin Content %%%%%%%
\input{tex/introduction}
\input{tex/research-method}
\input{tex/research-execution}
\input{tex/results}
\input{tex/discussion}
\input{tex/threats-to-validity}
\input{tex/conclusion}
\input{tex/data-availability-statements}
%%%%%%% End Content %%%%%%%%%

\begin{comment}

\section{Insert A head here}
This demo file is intended to serve as a ``starter file''
for IMAIAI journal papers produced under \LaTeX\ using
imaiai.cls v1.5e.

\subsection{Insert B head here}
Subsection text here.

\subsubsection{Insert C head here}
Subsubsection text here.

\section{Equations}

Sample equations.

%%% Numbered equation
\begin{align}\label{1.1}
\begin{split}
\frac{\partial u(t,x)}{\partial t} &= Au(t,x) \left(1-\frac{u(t,x)}{K}\right)-B\frac{u(t-\tau,x) w(t,x)}{1+Eu(t-\tau,x)},\\
\frac{\partial w(t,x)}{\partial t} &=\delta \frac{\partial^2w(t,x)}{\partial x^2}-Cw(t,x)+D\frac{u(t-\tau,x)w(t,x)}{1+Eu(t-\tau,x)},
\end{split}
\end{align}

\begin{align}\label{1.2}
\begin{split}
\frac{dU}{dt} &=\alpha U(t)(\gamma -U(t))-\frac{U(t-\tau)W(t)}{1+U(t-\tau)},\\
\frac{dW}{dt} &=-W(t)+\beta\frac{U(t-\tau)W(t)}{1+U(t-\tau)}.
\end{split}
\end{align}

%%%% Unnumbered equation
\[
\frac{\partial(F_1,F_2)}{\partial(c,\omega)}_{(c_0,\omega_0)} = \left|
\begin{array}{ll}
\frac{\partial F_1}{\partial c} &\frac{\partial F_1}{\partial \omega} \\\noalign{\vskip3pt}
\frac{\partial F_2}{\partial c}&\frac{\partial F_2}{\partial \omega}
\end{array}\right|_{(c_0,\omega_0)}
\]

\section{Enunciations}
%%%% Most of the enunciations like theorem, lemma, corollary, proposition, defintion,
%%%% condition, example, conjecture etc. are defined in the class file.

%%%% If the author wants to add or modify the enunciation style
%%%% they can define in the preamble as shown below.

%%%% \newtheoremstyle{theorem}{6pt}{6pt}{\rm}{}{\sffamily}{ }{ }{}
%%%% \theoremstyle{theorem}
%%%% \newtheorem{theorem}{\sc Theorem}[section]

%%%%\newtheoremstyle{corollary}{6pt}{6pt}{\rm}{}{\sffamily}{ }{ }{}
%%%%\theoremstyle{corollary}
%%%%\newtheorem{corollary}{\sc Corollary}[section]

%%%%\newtheoremstyle{definition}{6pt}{6pt}{\rm}{}{\sffamily}{ }{ }{}
%%%%\theoremstyle{definition}
%%%%\newtheorem{definition}[theorem]{\sc Definition}
%%%%
%%%%\newtheorem{exercise}[theorem]{Exercise}

\begin{theorem}\label{T0.1}
Assume that $\alpha>0, \gamma>1, \beta>\frac{\gamma+1}{\gamma-1}$.
Then there exists a small $\tau_1>0$, such that for $\tau\in
[0,\tau_1)$, if $c$ crosses $c(\tau)$ from the direction of
to  a small amplitude periodic traveling wave solution of
(2.1), and the period of $(\check{u}^p(s),\check{w}^p(s))$ is
\[
\check{T}(c)=c\cdot \left[\frac{2\pi}{\omega(\tau)}+O(c-c(\tau))\right].
\]
\end{theorem}

\section{Figures \& Tables}

The output for figure is:

\begin{figure}[!h]
%\centering\includegraphics[width=2.5in]{figurename.eps}
%%%call your figure name in the place "figurename.eps"
\caption{Insert figure caption here}
\label{fig_sim}
\end{figure}

 An example of a double column floating figure using two subfigures.
 (The subfig.sty package must be loaded for this to work.)
 The subfigure \verb+\label+ commands are set within each subfloat command, the
 \verb+\label+ for the overall figure must come after \verb+\caption+.
 \verb+\hfil+ must be used as a separator to get equal spacing.
 The subfigure.sty package works much the same way, except \verb+\subfigure+ is
 used instead of \verb+\subfloat+.

%\begin{figure*}[!h]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{figurename.eps}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{figurename.eps}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}

\vskip2pc

\noindent The output for table is:

\begin{table}[!h]
\caption{An Example of a Table}%%%Table caption goes here
\label{table_example}
\centering
\begin{tabular}{|c||c|}%%%The number of columns has to be defined here
\hline
One & Two\\ %%%% Table body
\hline
Three & Four\\%%%% Table body
\hline
\end{tabular}
\end{table}%%%End of the table

\section{Conclusion}
The conclusion text goes here.
\end{comment}

\section*{Acknowledgment}

Insert the Acknowledgment text here.

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/

\bibliographystyle{imaiai}
\bibliography{bibliography}
%
% once the .bbl file has been generated then place the text in your article.

% To get the numbered reference style the author should use [numbib]
%as an option in the document class.  For example: \documentclass[numbib]{imaiai}
\begin{comment}

\begin{thebibliography}{9}
\bibitem{1}
Vicsek T., Czir\'{o}k A., Ben-Jacob E., Cohen I.: `Novel type of phase transition in a system of self-driven particles', \textit{Phys. Rev. Lett.}, 1995, \textbf{75}, pp.~1226--1229

\bibitem{2}
Jadbabaie A., Lin J., Morse A.S.: `Coordination of groups of
mobile autonomous agents using nearest neighbor rules', \textit{IEEE
Trans. Autom. Control}, 2003, \textbf{48}, pp.~988--1001

\bibitem{3}
Olfati-Saber R., Murray R.M.: `Consensus problems in networks
of agents with switching topology and time-delays', \textit{IEEE Trans.
Autom. Control}, 2004, \textbf{49}, pp.~1520--1533
\end{thebibliography}
\end{comment}

\end{document}