### 2024-01-03 Article.tex

%% Author_tex.tex
%% V1.0
%% 2016/09/02
%% developed by Techset
%%
%% This file describes the coding for cta-author.cls

\documentclass{cta-author}%%%%where cta-author is the template name

\newtheorem{theorem}{Theorem}{}
\newtheorem{corollary}{Corollary}{}
\newtheorem{remark}{Remark}{}

%%%%%%Begin MyPackages%%
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{color}
\usepackage{tabulary}
\usepackage{enumitem}
	\setlist[itemize]{leftmargin=5.5mm}

%\usepackage{hyperref}
%\usepackage{url}
\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
\hypersetup{breaklinks=true}
\urlstyle{same}

\usepackage{natbib}
\setcitestyle{square}

\usepackage{soul}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{caption}
\usepackage[title]{appendix}
\usepackage{ulem}
\usepackage{placeins}
\usepackage{balance}
\usepackage[autostyle, english = american]{csquotes}
\MakeOuterQuote{"}

\usepackage[draft]{changes}

%%%%%%end MyPackages%%%%

%% Para arreglar las viñetas

\setitemize{label={\raisebox{0pt}{\normalsize\textbullet}}, labelindent=0em,labelsep=1em,leftmargin=*}

%% Hay otros cambios en el .cls

\begin{document}

\supertitle{Submission Template for IET Research Journal Papers}

\title{Investigation of the activities performed by experimental researchers in a software engineering lab using an ethnographic methodology}

\author{\au{Efraín R. Fonseca C.$^{1\corr}$}, \au{Oscar Dieste$^{2}$}, \au{Natalia Juristo$^{2}$}}

\address{\add{1}{Departamento de Ciencias de la Computación, Universidad de las Fuerzas Armadas - ESPE, Av. General Rumiñahui s/n y Ambato, Quito, Ecuador}
\add{2}{Departamento de Lenguajes y Sistemas Informáticos e Ingeniería de Software, Universidad Politécnica de Madrid, Calle de los Ciruelos, 28660, Madrid, España}
%\add{3}{Third Department, Third University, Address, Country Name}
%\add{4}{Current affiliation: Fourth Department, Fourth University, Address, Country Name}
\email{erfonseca@espe.edu.ec}}

\input{tex/abstract}

\maketitle

%%%%%%% Begin Content %%%%%%%
\input{tex/introduction}
\input{tex/research-method}
\input{tex/research-execution}
\input{tex/results}
\input{tex/discussion}
\input{tex/threats-to-validity}
\input{tex/conclusion}
%\input{tex/data-availability-statements}
%%%%%%% End Content %%%%%%%%%

\section*{Acknowledgment}
This work was supported by the project \replaced{PGC2018-097265-B-I00}{PID2022-137846NB-I00} funded by the Spanish Ministry of Science and Innovation and co-financed by FEDER, and the Universidad de las Fuerzas Armadas  - ESPE's Grupo de Investigación en Modelos de Producción de Software e Industria Inteligente (GrIMPSoftII). This work's preprint has been published \cite{Fonseca-2023-experimentation-proccess-SE-unpublished}.

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/

\Urlmuskip=0mu plus 1mu\relax
\bibliographystyle{iet}
\bibliography{bibliography}

\end{document}