### 2024-06-24 tex/appendix.tex

\appendices
\section{Acronyms}

\begin{table}[h]
	\caption{Acronyms used in the manuscript}
	\label{table}
	\setlength{\tabcolsep}{3pt}
	\begin{tabular}{|p{25pt}|p{115pt}|p{25pt}|}
	\hline
	Item & Term & Acronym\\
	\hline
	1 & Empirical Software Engineering & ESE\\
	\hline
	2 & Research Group Under Study & RGUS\\
	\hline
	3 & Grupo de Investigación en Modelos de Producción de Software e Industria Inteligente & GrIMPSoftII\\
	\hline
	4 & Research Manager & RM\\
	\hline
	5 & Experiment Manager & EM\\
	\hline
	6 & Senior Experimenter & SrEx\\
	\hline
	7 & Experimental Process Model & EPM\\
	\hline
	8 & Experimental Process Workflow & EPW\\
	\hline
	\multicolumn{3}{p{251pt}}{These acronyms have been used throughout the manuscript.}\\
	\end{tabular}
	\label{tab1}
\end{table}