### 2024-06-25 tex/appendix.tex

\appendices

\section{Acronyms}

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
\begin{table}[h]
\small
\center
\caption{Acronyms used throughout the manuscript}
\label{tab:acronyms}
\begin{tabular}{ll}
\hline
\multicolumn{1}{c}{Acronym} & \multicolumn{1}{c}{Term}       \\ \hline
\added[id=v4]{EM}                          & \added[id=v4]{Experiment Manager}             \\ 
\added[id=v4]{EPM}                         & \added[id=v4]{Experimental Process Model}     \\ 
\added[id=v4]{EPW}                         & \added[id=v4]{Experimental Process Workflow}  \\ 
\added[id=v4]{ESE}                         & \added[id=v4]{Empirical Software Engineering} \\ 
\added[id=v4]{RGUS}                        & \added[id=v4]{Research Group Under Study}     \\ 
\added[id=v4]{RM}                          & \added[id=v4]{Research Manager}               \\ 
\added[id=v4]{SE}                          & \added[id=v4]{Software Engineering}           \\ 
\added[id=v4]{SrEx}                        & \added[id=v4]{Senior Experimenter}            \\ \hline 
\end{tabular}
\end{table}